A fully labelled proof system for intuitionistic modal logics

### 2019-10-08 caleidoscope/soundness.tex
1 answer
rewrote section 7

\begin{proof}
	We proceed by induction on the height of the derivation of $\SG$, and we show for all rules in $\labIKp$
	$$
	\vliiinf{\labrn{r}}{}{\SG}{\SGi1}{\cdots}{\SGi n}
	$$ for $n\in\{0,1,2\}$, that whenever $\SGi1,\ldots,\SGi n$ are
	valid in all birelational frame, then so is $\SG$. It follows a case analysis on $\rn r$:
	\begin{itemize}
		\item $\rlabrn\bot$: This is trivial because $\bot$ is never forced.
		\item $\rn{id}$: This follows immediately from Proposition~\ref{prop:mon-adm}.
		\item $\llabrn\BOX$: By way of contradiction, assume that $\B, \Left,
		\futs xy, \accs yz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right$
		is valid in all birelational frames, but $\B,\Left, \futs xy,
		\accs yz, \labels{x}{\BOX A} \SEQ \Right$ is not. This means that
		we have a model $\M$ and an interpretation $\inter\cdot$, such
		that $\M,\inter\cdot\not\Vdash\B,\Left, \futs xy, \accs yz,
		\labels{x}{\BOX A} \SEQ \Right$, i.e., $\inter x\le_\M\inter y$
		and $\inter y\rel_\M\inter z$ and $\cforce \M x{\BOX A}$ and
		$\cnforce \M wB$ for all $\labels wB\in\Right$. However, by the
		definition of forcing in~\eqref{eq:kripke} we also have $\cforce
		\M zA$, and consequently $\nforce \M{\inter\cdot}{\B, \Left, \futs
			xy, \accs yz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ
			\Right}$. Contradiction.
		\todo{The reasoning by contradiction is not necessary. Could be rewritten more lightly as follows:}
			%
			Assume that $\B, \Left, \futs xy, \accs yz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right$
			is valid in all birelational frames. 
			Consider a model $\M$ and an interpretation $\inter\cdot$ of the conclusion sequent
			$\B,\Left, \futs xy, \accs yz, \labels{x}{\BOX A} \SEQ \Right$ into $\M$, such
			that for all $\labels{z}{C} \in \Gamma$ we have $\M, \inter z \Vdash \fm C$
			and $\cforce \M x{\BOX A}$. %and $\cnforce \M wB$ for all $\labels wB\in\Right$. However, 
			Hence, as $\inter x\le_\M\inter y$ and $\inter y\rel_\M\inter z$, 
			by the definition of forcing in~\eqref{eq:kripke} we also have $\cforce \M zA$, and 
			consequently $\cforce \M wB$ for all $\labels wB\in\Right$. 
			Thus, $\cforce \M{\inter\cdot}{\B, \Left, \futs
				xy, \accs yz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ
				\Right}$. \todo{should we rewrite the whole proof?}

>>>>

\begin{proof}
	We proceed by induction on the height of the derivation of $\SG$, and we show for all rules in $\labIKp$
	$$
	\vliiinf{\labrn{r}}{}{\SG}{\SGi1}{\cdots}{\SGi n}
	$$ for $n\in\{0,1,2\}$, that whenever $\SGi1,\ldots,\SGi n$ are
	valid in all birelational frame, then so is $\SG$. It follows a case analysis on $\rn r$:
	\begin{itemize}
		\item $\rlabrn\bot$: This is trivial because $\bot$ is never forced.
		\item $\rn{id}$: This follows immediately from Proposition~\ref{prop:mon-adm}.
		\item $\llabrn\BOX$: By way of contradiction, assume that $\B, \Left,
		\futs xy, \accs yz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right$
		is valid in all birelational frames, but $\B,\Left, \futs xy,
		\accs yz, \labels{x}{\BOX A} \SEQ \Right$ is not. This means that
		we have a model $\M$ and an interpretation $\inter\cdot$, such
		that $\M,\inter\cdot\not\Vdash\B,\Left, \futs xy, \accs yz,
		\labels{x}{\BOX A} \SEQ \Right$, i.e., $\inter x\le_\M\inter y$
		and $\inter y\rel_\M\inter z$ and $\cforce \M x{\BOX A}$ and
		$\cnforce \M wB$ for all $\labels wB\in\Right$. However, by the
		definition of forcing in~\eqref{eq:kripke} we also have $\cforce
		\M zA$, and consequently $\nforce \M{\inter\cdot}{\B, \Left, \futs
			xy, \accs yz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ
			\Right}$. Contradiction.
%		\todo{The reasoning by contradiction is not necessary. Could be rewritten more lightly as follows:}
%			%
%			Assume that $\B, \Left, \futs xy, \accs yz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right$
%			is valid in all birelational frames. 
%			Consider a model $\M$ and an interpretation $\inter\cdot$ of the conclusion sequent
%			$\B,\Left, \futs xy, \accs yz, \labels{x}{\BOX A} \SEQ \Right$ into $\M$, such
%			that for all $\labels{z}{C} \in \Gamma$ we have $\M, \inter z \Vdash \fm C$
%			and $\cforce \M x{\BOX A}$. %and $\cnforce \M wB$ for all $\labels wB\in\Right$. However, 
%			Hence, as $\inter x\le_\M\inter y$ and $\inter y\rel_\M\inter z$, 
%			by the definition of forcing in~\eqref{eq:kripke} we also have $\cforce \M zA$, and 
%			consequently $\cforce \M wB$ for all $\labels wB\in\Right$. 
%			Thus, $\cforce \M{\inter\cdot}{\B, \Left, \futs
%				xy, \accs yz, \labels{x}{\BOX A}, \labels{z}{A} \SEQ
%				\Right}$. \todo{should we rewrite the whole proof?}\lutz{no. we do by contradiction. I'll explainon skype.}
